### 2020-07-29 draft.Rmd

---
header-includes:
   - \usepackage{ulem}
   - \usepackage{float}
   - \usepackage{algorithm}
   - \usepackage{algorithmic}
output:
  pdf_document
fontsize: 10pt 
editor_options: 
  chunk_output_type: console
bibliography: references.bib  
---

@@

```{r r-setup}
library(tidyverse)
library(haven) # reading Stata data
library(MASS) # M-estimation
library(Rfit) # rank-based regression
library(patchwork)
library(broom)
library(sandwich)
library(lmtest)
library(NSM3)

@@

The key outcomes of interest for this reanalysis, the change in balance after six months and the change in balance after twelve months, are both highly right skewed.

```{r histogram}
#
# Create a histogram to show the skew of balchange
#
skewa = six_months %>%
  ggplot(aes(balchange)) +
  geom_histogram() +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(limits = c(0, 2000),
                     labels = scales::comma) +
  labs(title = "The Dependent Variables Are Highly Right Skewed",
       subtitle = "Six Month Change in Balance",
       x = "Change in balance (Philippine Peso)",       
       y = NULL) +
  theme_minimal()

@@

Here is the same information as a kernel density estimate (KDE).

```{r density-plot}
#
# Create a density plot to show the skew of balchange
#
skewa = six_months %>%
  ggplot(aes(balchange)) +
  geom_density() +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(limits = c(0, 0.004)) +
  labs(title = "The Dependent Variables Are Highly Right Skewed",
       subtitle = "Six Month Change in Balance",
       x = "Change in balance (Philippine Peso)",       
       y = NULL) +
  theme_minimal()

@@

The right skew is present in the treatment and control groups, but it most pronounced in the commit-savings treatment group after twelve months.

```{r skew-by-group}
#
# Create a plot to show the skew of balchange by group
#
skewa = six_months %>%
  ggplot(aes(balchange, y = group, color = group)) +
  geom_point(alpha = 0.2) +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "The Dependent Variables Are Highly Right Skewed",
       subtitle = "Six Month Change in Balance",
       x = "Change in balance (Philippine Peso)",       
       y = NULL) +
  theme_minimal() +
  guides(color = FALSE)

@@

@@

```{r wilcox-paired-test}
wilcox.test(seed_accounts$totbal, seed_accounts$newtotbal,alternative = "greater", paired=TRUE)

@@

```{r 2-sample}
x = list(nonseed$balchange, seed_accounts$balchange)
fp_result = pFligPoli(x, method="Monte Carlo")
fp_result
```

@@

```{r Tukey-HSD}

seed_balchange = seed_accounts$balchange
marketing_balchange = treated %>% 
  filter(marketing==1)%>%
  pull(balchange)
control_balchange = treated %>% 
  filter(control==1, marketing==0) %>%
  pull(balchange)

@@

kw_result = kruskal.test(treat_list)
kw_result

@@

for (row in names(treat_list)){
  for(col in names(treat_list)){
    m[row,col] = wilcox.test(
      treat_list[[row]],
      treat_list[[col]],
      alternative="greater")$p.value
  }
}
#print("pairwise statistics greater than or equal to the adjusted alpha critical value")
comparison_matrix = as.table(m <= adjusted_alpha)
knitr::kable(comparison_matrix, caption="Significant difference found in pairwise comparison with Wilcoxon Test.")

@@

The dependent variable, change in savings account balance, is highly right-skewed after six months and twelve months as demonstrated in [todo(aaron): add plot number from above]. Accordingly, the dependent variable is not approximately normally distributed and the error term of their estimated models is not approximately normally distributed.

[todo(aaron): add plot number from above] The SEED treatment, the marketing control, and the full control all have outliers. Though the SEED treatment's outliers are distinct for the change in twelve month balance.

@@

count(treated, change) %>%
  pivot_wider(names_from = change,
              values_from = n) %>%
  knitr::kable(caption = "Change in Balance After Twelve Months")

@@

```{r blachange-direction-by-group}
count(treated, group, change) %>%
  pivot_wider(names_from = change,
              values_from = n) %>%
  knitr::kable(caption = "Change in Balance After Twelve Months by Treatment Group")

@@

```{r porbit-dependent-vars}
count(treated, frac_change_20) %>%
  pivot_wider(names_from = frac_change_20,
              values_from = n) %>%
  rename(`20%+ Increase` = `1`,
         `<20% Increase` = `0`) %>%
  knitr::kable(caption = "20+% Change in Balance After Twelve Months")

count(treated, group, frac_change_20) %>%
  pivot_wider(names_from = frac_change_20,
              values_from = n) %>%  
  rename(`20%+ Increase` = `1`,
         `<20% Increase` = `0`) %>%  
  knitr::kable(caption = "20+% Change in Balance After Twelve Months")

@@

We estimate identical specifications to [@Ashraf2006] using M-estimation. Column (1) shows the change in balance after six months including all observations. The coefficients estimate the additional savings of the corresponding treatment over the control group. Column (2) shows the change in balance after six months using only the commitment savings group and the marketing group. Column (3) is similar to column (1), only it uses the twelve month change in balance. Likewise, column (4) is similar to column (2), only it uses the twelve month change in balance.

@@

We estimate identical specifications to [@Ashraf2006] using robust and rank-based regressions. Column (1) shows the change in balance after six months including all observations. The coefficients estimate the additional savings of the corresponding treatment over the control group. Column (2) shows the change in balance after six months using only the commitment savings group and the marketing group. Column (3) is similar to column (1), only it uses the twelve month change in balance. Likewise, column (4) is similar to column (2), only it uses the twelve month change in balance.

@@

The models estimated with OLS have residuals that are highly skewed. The following QQ plots demonstrate that many of the residuals are approximately normally distributed, as they fall on the diagonal line. However, the residuals dramatically depart the line at higher values.

```{r residual-plot}
resid1 = augment(lm3) %>%
  ggplot(aes(sample = .resid)) +
  geom_qq(alpha = 0.2) +
  geom_qq_line() +
  theme_minimal() +
  labs(title = "The Estimated Model has Significant Outliers",
       subtitle = "QQ Plot for OLS (3)")

@@

The following data  compares the estimated coefficients and confidence intervals from column (3) for all three model types.

@@

coefficients %>%
  mutate(conf.low = estimate - qnorm(0.025) * std.error,
         conf.high = estimate + qnorm(0.025) * std.error) %>%
  mutate(model = factor(model, levels = c("Rank-Based\nModel", "M Model", "OLS Model"))) %>%
  ggplot(aes(x = estimate, 
             y = model,
             xmin = conf.low,
             xmax = conf.high,
             color = model)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  facet_wrap(~term, scales = "free_x") +
  labs(title = "Model Choice Affects Estimated Coefficients",
       x = "Estimated Beta and 95% Confidence Interval",
       y = NULL) +
  theme_bw() +
  theme(legend.position = "top")

@@

The following data visualization compares the estimated coefficients and confidence intervals from column (4) for all three model types.

@@

coefficients %>%
  mutate(conf.low = estimate - qnorm(0.025) * std.error,
         conf.high = estimate + qnorm(0.025) * std.error) %>%
  mutate(model = factor(model, levels = c("Rank-Based\nModel", "M Model", "OLS Model"))) %>%
  ggplot(aes(x = estimate, 
             y = model,
             xmin = conf.low,
             xmax = conf.high,
             color = model)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  facet_wrap(~term, scales = "free_x") +
  labs(title = "Model Choice Affects Estimated Coefficients",
       x = "Estimated Beta and 95% Confidence Interval",
       y = NULL) +
  theme_bw() +
  theme(legend.position = "top")

>>>>

---
header-includes:
   - \usepackage{ulem}
   - \usepackage{float}
   - \usepackage{algorithm}
   - \usepackage{algorithmic}
output:
  pdf_document
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
bibliography: references.bib  
---

@@

```{r r-setup}
library(tidyverse)
library(haven) # reading Stata data
library(MASS) # M-estimation
library(Rfit) # rank-based regression
library(patchwork) # for combining ggplot2s
library(broom) # for tidying regression output
library(sandwich) # for HC1 robust standard errors
library(lmtest) # for HC1 robust standard errors
library(NSM3)
library(knitr) # for tables

@@

The key outcomes of interest for this reanalysis, the change in balance after six months and the change in balance after twelve months, are both highly right skewed as demonstrated in \textit{figure 1}.

```{r histogram}
#
# Create a histogram to show the skew of balchange
#
skewa = six_months %>%
  ggplot(aes(balchange)) +
  geom_histogram() +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(limits = c(0, 2000),
                     labels = scales::comma) +
  labs(title = "Figure 1: The Dependent Variables Are Highly Right Skewed",
       subtitle = "Six Month Change in Balance",
       x = "Change in balance (Philippine Peso)",       
       y = NULL) +
  theme_minimal()

@@

\textit{Figure 2} displays the same data as \textit{figure 1} as a kernel density estimate (KDE).

```{r density-plot}
#
# Create a density plot to show the skew of balchange
#
skewa = six_months %>%
  ggplot(aes(balchange)) +
  geom_density() +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(limits = c(0, 0.004)) +
  labs(title = "Figure 2: The Dependent Variables Are Highly Right Skewed",
       subtitle = "Six Month Change in Balance",
       x = "Change in balance (Philippine Peso)",       
       y = NULL) +
  theme_minimal()

@@

The right skew is present in the treatment and control groups, but it most pronounced in the commit-savings treatment group after twelve months as demonstrated in \textit{figure 3}.

```{r skew-by-group}
#
# Create a plot to show the skew of balchange by group
#
skewa = six_months %>%
  ggplot(aes(balchange, y = group, color = group)) +
  geom_point(alpha = 0.2) +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "Figure 3: The Dependent Variables Are Highly Right Skewed",
       subtitle = "Six Month Change in Balance",
       x = "Change in balance (Philippine Peso)",       
       y = NULL) +
  theme_minimal() +
  guides(color = FALSE)

@@

```{r variance-by-group-quantile}

skewc = treated %>%
  group_by(group) %>%
  filter(balchange < quantile(balchange, 0.99)) %>%
  ggplot(aes(balchange, y = group, color = group)) +
  geom_point(alpha=0.2) +
  scale_x_continuous(labels = scales::comma) +
  labs(title = "Figure 4: Variance by treatment having removed significant outliers",
       x = "Change in balance (Philippine Peso)",
       y = NULL) +  
  theme_minimal()

skewc
rm(skewc)

@@

```{r wilcox-paired-test}
wilcox_test = wilcox.test(seed_accounts$totbal,
                          seed_accounts$newtotbal,
                          alternative = "greater", 
                          paired = TRUE)

tibble(
  Metric = c("Test Statistic", "P-Value"),
  Value = c(wilcox_test$statistic, wilcox_test$p.value)
) %>%
  kable(caption = "Wilcoxon Signed-Rank Test Results")

@@

```{r 2-sample}
#
# Fligner-Policello test
#
x = list(nonseed$balchange, seed_accounts$balchange)
fp_result = pFligPoli(x, method="Monte Carlo")

tibble(
  Metric = c("Fligner-Policello U Statistic", "two-sided p-value"),
  Value = c(fp_result$obs.stat, fp_result$two.sided)
) %>%
  kable(caption = "Fligner-Policello Test Results")

Test conducted with `r fp_result$m` X values, `r fp_result$n` Y values, and 10000 iterations.

@@

```{r kruskal-wallis}
#
# Kruskal-Wallis test
#

seed_balchange = seed_accounts$balchange

marketing_balchange = treated %>% 
  filter(marketing==1)%>%
  pull(balchange)

control_balchange = treated %>% 
  filter(control==1, marketing==0) %>%
  pull(balchange)

@@

kw_result = kruskal.test(treat_list)

tibble(
  Metric = c("Chi-Squared", "Degrees Freedom", "P-Value"),
  Value = c(kw_result$statistic, kw_result$parameter, kw_result$p.value)
) %>%
  kable(caption = "Kruskal-Wallis Test Results")

@@

for (row in names(treat_list)){
  for(col in names(treat_list)){
    m[row,col] = wilcox.test(
      treat_list[[row]],
      treat_list[[col]],
      alternative="greater")$p.value
  }
}

as.table(m <= adjusted_alpha) %>%
  kable(caption = "Significant difference found in\n
        pairwise comparison with Wilcoxon Test.")

@@

The dependent variable, change in savings account balance, is highly right-skewed after six months and twelve months as demonstrated in \textit{figure 2}. Accordingly, the dependent variable is not approximately normally distributed and the error term of their estimated models is not approximately normally distributed.

As \textit{figure 3} demonstrated, the SEED treatment, the marketing control, and the full control all have outliers. Though the SEED treatment's outliers are distinct for the change in twelve month balance.

@@

count(treated, change) %>%
  pivot_wider(names_from = change,
              values_from = n) %>%
  kable(caption = "Change in Balance After Twelve Months")

@@

```{r blachange-direction-by-group}
count(treated, group, change) %>%
  pivot_wider(names_from = change,
              values_from = n) %>%
  kable(caption = "Change in Balance After Twelve Months by Treatment Group")

@@

```{r porbit-dependent-vars}
count(treated, frac_change_20) %>%
  pivot_wider(names_from = frac_change_20,
              values_from = n) %>%
  rename(`20%+ Increase` = `1`,
         `<20% Increase` = `0`) %>%
  kable(caption = "20+% Change in Balance After Twelve Months")

count(treated, group, frac_change_20) %>%
  pivot_wider(names_from = frac_change_20,
              values_from = n) %>%  
  rename(`20%+ Increase` = `1`,
         `<20% Increase` = `0`) %>%  
  kable(caption = "20+% Change in Balance After Twelve Months")

@@

We estimate identical specifications to [@Ashraf2006] using M-estimation. Column (1) in \textit{table 10} shows the change in balance after six months including all observations. The coefficients estimate the additional savings of the corresponding treatment over the control group. Column (2) shows the change in balance after six months using only the commitment savings group and the marketing group. Column (3) is similar to column (1), only it uses the twelve month change in balance. Likewise, column (4) is similar to column (2), only it uses the twelve month change in balance.

@@

We estimate identical specifications to [@Ashraf2006] using robust and rank-based regressions. Column (1) in \textit{table 11} shows the change in balance after six months including all observations. The coefficients estimate the additional savings of the corresponding treatment over the control group. Column (2) shows the change in balance after six months using only the commitment savings group and the marketing group. Column (3) is similar to column (1), only it uses the twelve month change in balance. Likewise, column (4) is similar to column (2), only it uses the twelve month change in balance.

@@

The models estimated with OLS have residuals that are highly skewed. \textit{Figure 5} displays QQ plots that demonstrate that many of the residuals are approximately normally distributed, as they fall on the diagonal line. However, the residuals dramatically depart the line at higher values.

```{r residual-plot}
resid1 = augment(lm3) %>%
  ggplot(aes(sample = .resid)) +
  geom_qq(alpha = 0.2) +
  geom_qq_line() +
  theme_minimal() +
  labs(title = "Figure 5: The Estimated Model has Significant Outliers",
       subtitle = "QQ Plot for OLS (3)")

@@

\textit{Figure 6} compares the estimated coefficients and confidence intervals from column (3) for all three model types.

@@

coefficients %>%
  mutate(conf.low = estimate - qnorm(0.025) * std.error,
         conf.high = estimate + qnorm(0.025) * std.error) %>%
  mutate(model = factor(model, 
                        levels = c("Rank-Based\nModel", "M Model", "OLS Model"))) %>%
  ggplot(aes(x = estimate, 
             y = model,
             xmin = conf.low,
             xmax = conf.high,
             color = model)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  facet_wrap(~term, scales = "free_x") +
  labs(title = "Figure 6: Model Choice Affects Estimated Coefficients",
       x = "Estimated Beta and 95% Confidence Interval",
       y = NULL) +
  theme_bw() +
  theme(legend.position = "top")

@@

\textit{Figure 7} compares the estimated coefficients and confidence intervals from column (4) for all three model types.

@@

coefficients %>%
  mutate(conf.low = estimate - qnorm(0.025) * std.error,
         conf.high = estimate + qnorm(0.025) * std.error) %>%
  mutate(model = factor(model, 
                        levels = c("Rank-Based\nModel", "M Model", "OLS Model"))) %>%
  ggplot(aes(x = estimate, 
             y = model,
             xmin = conf.low,
             xmax = conf.high,
             color = model)) +
  geom_pointrange() +
  geom_vline(xintercept = 0, alpha = 0.5) +
  facet_wrap(~term, scales = "free_x") +
  labs(title = "Figure 7: Model Choice Affects Estimated Coefficients",
       x = "Estimated Beta and 95% Confidence Interval",
       y = NULL) +
  theme_bw() +
  theme(legend.position = "top")
